---
title: "Tidy Data/Data Relations"
author: "Me"
date: "28/08/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
library(tidyverse) # loads packages for data manipulation (dplyr/tidyr)
```

## Resources

* [Tidy Data](https://psyteachr.github.io/msc-data-skills/tidyr.html) in Data Skills for Reproducible Science
* [Data Relations](https://psyteachr.github.io/msc-data-skills/joins.html) in Data Skills for Reproducible Science
* [Data Transformation Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
* [Stub script for this lesson](stubs/1_tidy.Rmd)

## Example

### Load Data

```{r}
disgust <- read_csv("data/disgust.csv")

users <- read_csv("data/users.csv")

```


### Merge tables

Write my notes here. Merge disgust and user tables to one table.

```{r}
joined_data <- left_join(disgust, users, by = "user_id")
```

### Filter

Select only rows where sex is either "male" or "female"

```{r}
filtered_data <- filter(joined_data, sex %in% c("male", "female"))
```

### Wide to long format

Transform from 1 row per subject and 21 questions per row, to one question:subject per row

```{r}
long_data <- gather(filtered_data, 
                    key = question, 
                    value = score, 
                    moral1:pathogen7)
```

### Separate a column

The new "question" column is n the format "domain#". Split this into two columns for "domain" and "n".

```{r}
separated_data <- separate(long_data,
                           col = question,
                           into = c("domain", "q_n"),
                           sep = -1)
```

### Summarise

Group by domain and user_id to calculate the sum of the 7 items within each domain for each subject.

```{r}
grouped_data <- group_by(separated_data, user_id, sex, domain)

aggregated_data <- summarise(
  grouped_data, 
  mean_score = mean(score)
)
```

### Long to Wide

```{r}
wide_data <- spread(aggregated_data, domain, mean_score)

```

Now, save your script, restart R, and knit (or run) the whole script!

### The Pipe

Or you can do this all in one step by [piping](https://psyteachr.github.io/msc-data-skills/tidyr.html#pipes). The pipe sends the results of the last function to the first argument of the next function. Remember to remove the first argument of the function when converting from separate functions to piped functions.

```{r}
data <- left_join(disgust, users, by = "user_id") %>%
  filter(sex %in% c("male","female")) %>%
  gather(question, score, moral1:pathogen7) %>%
  separate(question, c("domain", "q_n"), sep = -1) %>%
  group_by(user_id, sex, domain) %>%
  summarise(mean_score = mean(score)) %>%
  spread(domain, mean_score) %>%
  select(user_id, moral:sexual, user_sex = sex)

```

## Tidying Functions

```{r}
# load the data
ocean <- read_csv("data/personality.csv")
```


### gather()

```{r}
ocean_gathered <- gather(ocean, question, score, Op1:Ex9)

head(ocean_gathered)
```

### spread()

```{r}
ocean_spread <- spread(ocean_gathered, question, score)
```

### separate()

```{r}

ocean_sep <- separate(ocean_gathered, question, c("domain", "n"), sep = 2) %>%
  separate(date, c("year", "month", "day"), sep = "-")

```



### unite()

Convert year, month and day columns into a new column called `date` with the format "31/02/2019"

```{r}
ocean_unite <- unite(ocean_sep, date, day, month, year, sep = "/")
```



## Data Relations

Below we create two small tables to practice joining data.

```{r}
subject <- tibble(
  id = seq(1,5),
  sex = c("m", "m", NA, "f", "f"),
  age = c(19, 22, NA, 19, 18)
)

exp <- tibble(
  id = c(2, 3, 4, 4, 5, 5, 6, 6, 7),
  score = c(10, 18, 21, 23, 9, 11, 11, 12, 3)
)
```


### left_join()

```{r}
left_join(exp, subject, by = "id")

left_join(subject, exp, by = "id")
```

### right_join()

Just a backwards left join. Only changes column order. Very seldom used.

```{r}
right_join(exp, subject, by = "id")

right_join(subject, exp, by = "id")
```

### inner_join()

```{r}
inner_join(exp, subject, by = "id")

inner_join(subject, exp, by = "id")
```

### full_join()

```{r}
full_join(exp, subject, by = "id")

full_join(subject, exp, by = "id")
```

### semi_join()

```{r}
semi_join(exp, subject, by = "id")

semi_join(subject, exp, by = "id")
```

### anti_join()

```{r}
anti_join(exp, subject, by = "id")

anti_join(subject, exp, by = "id")
```

