---
title: "Data Visualisation"
author: "Lisa"
date: "29/08/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
library(tidyverse) # loads packages for data manipulation (dplyr/tidyr) and visualisation (ggplot2)
library(plotly)
library(cowplot)

theme_set(theme_minimal()) # set default theme for plots
```

## Resources

* [Data Visualisation](https://psyteachr.github.io/msc-data-skills/ggplot.html) in Data Skills for Reproducible Science
* [Data Visualisation Cheatsheet](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)
* [Stub script for this lesson](stubs/3_viz.Rmd)
* [Plot Demo Shiny App](http://shiny.psy.gla.ac.uk/debruine/plotdemo/)


The code below creates a few data files for our examples.

```{r}
pets <- tibble(
  pet = sample(
    c("dog", "cat", "ferret", "bird", "fish"), 
    100, 
    TRUE, 
    c(0.45, 0.40, 0.05, 0.05, 0.05)
  )
)

pet_happy <- tibble(
  pet = rep(c("dog", "cat"), each = 500),
  happiness = c(rnorm(500, 55, 10), rnorm(500, 45, 10)),
  age = rpois(1000, 3) + 20
)

x_vs_y <- tibble(
  x = rnorm(100),
  y = x + rnorm(100, 0, 0.5)
)

overlap <- tibble(
  x = rbinom(1000, 10, 0.5),
  y = x + rbinom(1000, 20, 0.5)
)

overplot <- tibble(
  x = rnorm(10000),
  y = x + rnorm(10000, 0, 0.5)
)
```


## geom_bar()

```{r}
ggplot(pets, aes(x = pet, fill = pet)) +
  geom_bar(color = "grey", size = 2, show.legend = FALSE)
```

## geom_histogram()

```{r}
ggplot(pet_happy, aes(x = age)) +
  geom_histogram(binwidth = 1,
                 fill = "white",
                 color = "grey")
```

## geom_density()

```{r}
ggplot(x_vs_y) +
  geom_density(aes(x = x), color = "red") +
  geom_density(aes(x = y), color = "blue")
```


```{r}
x_vs_y %>%
  gather(xy, value, x, y) %>%
  ggplot(aes(x = value, fill = xy)) +
  geom_density(alpha = 0.5)
```

## geom_freqpoly()

```{r, fig.width = 2, fig.height=3, fig.align="center", fig.cap="histogram plus freqpoly"}
ggplot(pet_happy, aes(x = age)) +
  geom_histogram(binwidth = 1,
                 fill = "white",
                 color = "grey") +
  geom_freqpoly(binwidth = 1)
```



## geom_violin()

```{r}
ggplot(pet_happy, aes(x = pet, y = happiness, fill = pet, color = pet)) +
  geom_violin() +
  scale_fill_manual(values = c("green", "purple"),
                    labels = c("Cats", "Dogs"),
                    name = "Type of Pet") +
  scale_color_manual(values = c("palevioletred2", "dodgerblue"),
                    labels = c("Cats", "Dogs"),
                    name = "Type of Pet")
```

## geom_boxplot()

```{r}
ggplot(pet_happy, aes(x = pet, y = happiness, color = pet)) +
  geom_violin(show.legend = FALSE) +
  geom_boxplot(aes(fill = pet), color = "grey30", width = 0.3) +
  coord_flip()
```


## geom_point()

```{r}
ggplot(x_vs_y, aes(x = x, y = y)) +
  geom_point(shape = 0) +
  labs(x = "Sunshine",
       y = "Happiness",
       title = "Sunshine vs Happiness")
```

## geom_smooth()

```{r}
g <- ggplot(pet_happy, aes(x = age, y = happiness, 
                      color = pet, linetype = pet)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE)

ggplotly(g)
```

```{r}
ggplot(overlap, aes(x, y)) +
  geom_point(size = 5, alpha = .05)
```

```{r}
ggplot(overlap, aes(x, y)) +
  geom_bin2d(binwidth = 1, drop = FALSE) +
  scale_fill_viridis_c(option = "viridis")
```

```{r}
ggplot(overlap, aes(x, y)) +
  geom_count()
```

```{r}
ggplot(overplot, aes(x, y)) +
  geom_density_2d()
```

```{r}
ggplot(overplot, aes(x, y)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon") +
  scale_fill_viridis_c()
```


## Multiple Plots

Make each individual plot

```{r}
g <- ggplot(pet_happy, aes(x = pet, y = happiness, fill = pet)) +
  scale_fill_manual(values = c("green", "purple"))

boxplot <- g + geom_boxplot(show.legend = FALSE)

violinplot <- g + geom_violin(show.legend = FALSE)

violinbox <- g + geom_violin(show.legend = FALSE, fill = "white") +
  geom_boxplot(show.legend = FALSE, width = 0.3)

ggsave("images/boxplot.png", boxplot, width = 5, height = 3, dpi = 300)

```


```{r}

cowplot::plot_grid(boxplot, violinplot, violinbox, 
                   nrow = 1,
                   labels = c("A", "B", "C")
                   )

ggsave("images/cowplot.jpg", 
       width = 5,
       height = 3, 
       dpi = 300)

```

